Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Scenario

### 2017-02-21 documents/TriangleScenarioArXiv/ts.tex
minor edits, spelling/phrasing etc

% === Revtex Declaration ===
\documentclass[aps, 10pt, english, twoside, pra, nofootinbib, tightenlines, longbibliography]{revtex4-1}

@@

\begin{document}
    \title{Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Scenario}
    \author{Thomas C. Fraser}
    \email{tcfraser@tcfraser.com}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada \\ University of Waterloo, Waterloo, Ontario, Canada}
    % \author{Elie Wolfe}
    % \email{ewolfe@perimeter@institute.ca}
    % \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada}
    \date{\today}
    \begin{abstract}
        Quantum correlations are often incompatible with a classical assumption of causal structure. This non-classicality is often known as quantum non-locality, and it is witnessed through the violation of causal compatibility inequalities, such as Bell inequalities. Such inequalities were recently derived for the triangle scenario [arXiv:1609.00672], begging the question: can these inequalities be violated by quantum correlations? Here we answer this affirmatively, and discuss specific triangle scenario inequalities and quantum configurations which manifest non-classical correlations. Numerical optimizations reveal quantum resources potentially qualitatively different from those known previously.
    \end{abstract}
    \maketitle
    \tableofcontents
    \setlength{\parskip}{1em}
    \clearpage

    \section{Introduction}
    \label{sec:introduction}
    In recent decades, the technical utility of quantum mechanics has become abundantly clear; quantum computational algorithms such as Shor's algorithm~\cite{Shor_1997} and numerous others~\cite{Jordan_2016} maintain advantage over their classical counterparts. Additionally, classical communications technologies stand to benefit from the nature of quantum mechanics --- a popular example being Quantum Key Distribution~\cite{Bennett_2014}. Any time classical resources are unable to emulate quantum mechanics, there represents an opportunity to exploit the non-classical features of quantum mechanics in order to solve a variety of computational and communicational problems~\cite{Neilsen_Chaung_2011}. Albeit, not all tasks stand to benefit from quantum resources~\cite{Almeida_2010}. Presently, discovering situations or tasks where quantum mechanics offers an advantage, and proving that advantage is genuine, is a primal objective of theoretical quantum information theory.

    From a foundational prospective, the majority of attempts to characterize precisely where quantum mechanics and classical mechanics differ have involved Bell inequalities~\cite{Bell_1964}. Originally, Bell inequalities were derived in response to the famous EPR paradox~\cite{EPR_Orig} to rule out hidden variable theories for quantum mechanics. Studying and discovering Bell-type inequalities is useful because a complete set of Bell inequalities provides a systematic way of classifying the \textit{classicality} of a given observation. These inequalities are best understood from the perspective of \term{causal inference} --- a problem which aims to determine which observations can or can not be explained by a hypothesized causal structure. The abstract nature of causal inference is responsible for its presence in numerous scientific fields including machine learning and biology~\cite{Pearl_2009,Pearl_2009_tr}. Inequalities such as Bell inequalities constrain the space of observations that are compatible with a hypothesized causal structure. In the case of the original Bell inequalities, the classical causal model in question is known as the Bell scenario~\cite{Wood_2012} and is depicted in \cref{fig:bell_scenario}. The Bell scenario involves two parties ($A, B$) who each make measurements decided by random settings ($S_A, S_B$) on a shared hidden resource $\la$. Quantum non-classicality in the Bell scenario and related structures have been studied thoroughly since Bell's original work~\cite{Brunner_2013}. However, novel scenarios such as the correlation scenarios proposed by Fritz~\cite{Fritz_2012,Fritz_2014} are arguably less understood compared to the Bell scenario. Here we investigate one particular causal scenario: the triangle scenario (\cref{fig:triangle_scenario}).

    The triangle scenario (\cref{fig:triangle_scenario}) is a causal structure composed of $3$ parties labeled $A, B, C$ arranged in a triangular configuration while pair-wise sharing hidden/latent variables $X, Y, Z$. It has been studied extensively in literature before (see~\cite[Fig. 1]{Steudel_2010},~\cite[Fig. 6]{Chaves_2014},~\cite[Fig. 8]{Branciard_2012},~\cite[Fig. 8, Appendix E]{Henson_2014},~\cite[Fig. 3]{Fritz_2012},~\cite[Fig. 1]{Inflation}, $\ldots$). A healthy portion of which the existing literature will be discussed in \cref{sec:triangle_scenario}. For instance \citet{Branciard_2012} noted that characterizing locality in the triangle scenario remained an \textit{open problem} and that identifying causal compatibility inequalities for this configuration \textit{seemed challenging}. The subject of this document is to expound on the developments made towards understanding quantum non-classicality in the triangle scenario.

    As a preliminary summary, causal compatibility inequalities were derived for the triangle scenario that are able to be violated by specific quantum-accessible distributions. This was accomplished by combining two previous works. The first, an insight made by \citet{Fritz_2012} regarding the ability to re-interpret the Bell scenario as a portion of the triangle scenario. The second, a new framework for solving causal inference problems developed by Wolfe \emph{et al.}~\cite{Inflation} called \term{The Inflation Technique}. Ultimately, this work serves as a strong indication that the Inflation Technique offers new insights into specifically quantum non-classicality. Moreover, these new inequalities offer an avenue for finding and identifying new forms of non-classicality unknown previously. This report discusses attempts to find such resources with partial success and outlines refinements that can be made for future exploration.

    The following paper is organized into the following sections: \Cref{sec:causal_compatibility} recalls important notions from causal inference theory and sets up the notation to be used. \Cref{sec:triangle_scenario} discusses the triangle scenario and its existing research in detail; identifying its stark differences from the Bell scenario and motivates why the triangle scenario was chosen as a case study. \Cref{sec:fritz_distribution} defines and discusses a singular quantum correlation first discovered by \citet{Fritz_2012} which we term the \term{Fritz distribution}. The Fritz distribution was proven to be non-classical without the use of inequalities --- this report derives inequalities that do. \Cref{sec:inflation_technique} briefly summaries the Inflation Technique in the context of this work. Although the summary presented here is sufficient, the original work~\cite{Inflation} offers a more concrete and pedagogical introduction. \Cref{sec:deriving_inequalities} demonstrates how the inflation technique can be used to derive causal compatibility inequalities for the triangle scenario that are violated by quantum-accessible distributions. A sample of these inequalities are presented at the end of \cref{sec:deriving_inequalities} as well as \cref{sec:web_inequality,sec:symmetic_web_inequality}. Finally, \cref{sec:violations_noise} mentions a method for parameterizing the space of quantum-accessible distributions which is used to numerically optimize against our derived inequalities in search of stronger types of non-classicality. Additionally, the Fritz distribution is subjected to noise in order to measure the robustness of the derived inequalities.

    \section{Causal Compatibility}
    \label{sec:causal_compatibility}
    The task of causal inference aims to determine whether or not a given family of observed probability distributions can be explained by a proposed causal mechanism~\cite{Pearl_2009}. If a set of observed distributions can be explained by the hypothesized causal mechanism, then the observations are said to be \textit{compatible} with the causal mechanism. In order to define compatibility rigorously, we first need to formally define what is meant by ``set of observations''.

    \subsection{Marginal Models}
    Given a set of observable variables $\jointvar = \bc{v_1, \ldots, v_k}$ where each $v \in \jointvar$ is a statistical random variable, one can conceivably make a measurement over of these variables and obtain a joint probability distribution $\prob[\jointvar] = \prob[v_1,\ldots,v_n]$. More generally however, and depending on the context, it is possible to make a measurement on a portion of the system $V \subset \jointvar$ in order to obtain a marginal distribution $\prob[V]$. With complete generality however, one could in principle make a measurement on a collection of portions of the system $\mscenario = \bc{V_1, \ldots, V_m \mid \forall i : V_i \subseteq \jointvar}$ obtaining a collection of marginal probability distributions $\bc{\prob[V_1], \ldots, \prob[V_m] \mid \forall i: V_i \in \mscenario}$.

    The set $\mscenario = \bc{V \mid V \subseteq \jointvar}$ of subsets of $\jointvar$ is referred to as the \term{marginal scenario} and each element $V \in \mscenario$ is termed a \term{(marginal) context} of $\mscenario$. The complete set of marginal distributions is referred to as the \term{marginal model} and is denoted with an superscript $\prob^{\mscenario}$:
    \[ \prob^{\mscenario} = \bc{\prob[V] \mid V \in \mscenario} \]
    A marginal model acts as the most general description of a family of observations that can be made over $\jointvar$. Strictly speaking, as defined by~\cite{Fritz_2011}, a marginal scenario forms an \textit{abstract simplicial complex} where it is required that all subsets of contexts are also contexts.
    \[ \forall V \in \mscenario, V' \subset V : V' \in \mscenario \]
    Throughout this document, we only consider (without loss of generality) the \textit{maximal} marginal scenario; restricting our focus to the largest contexts in the marginal scenario. Additionally, all marginal scenarios are taken to be \text{complete} in the sense that the marginal scenario covers the complete set of observable variables $\jointvar$.
    \[ \jointvar = \bigcup_{V \in \mscenario} V \]

    \subsection{Causal Structure}
    A \term{causal structure} abstractly represents the hypothesized causal mechanism. Mathematically, a causal structure is a directed acyclic graph $\graph = \br{\nodes, \edges}$ of nodes $\nodes$ and edges $\edges$ as defined in \cref{sec:graph_theory}. The nodes of a causal structure are random variables and are classified into two disjoint classes: the \term{latent nodes} $\nodes_L$ representing the set of nodes that are unobservable and the \term{observed nodes} $\nodes_O$. The latent nodes of a causal structure represent completely hidden variables that are hidden by some fundamental process or cannot be measured due to other limitations. The observed nodes represent random variables that are measurable and in practice will be observed by the marginal model $\prob^\mscenario$. The edges $\edges$ of a causal structure represent casual influence between variables within the causal structure.

@@

    Now that the notions of a causal structure $\graph$ and a marginal model $\prob^{\mscenario}$ are defined, it is now possible to precisely define the notion of compatibility between the two. Intuitively, the causal connections within $\graph$ should restrict and limit the set of possible observations that can be made on $\graph$. Namely, if each variable $n \in \nodes$ is caused by its parents $\Pa[\graph]{n}$, then $n$ should be independent of all other variables that are not its descendants. This is known as the \term{local Markov property}:
    \[ n \indep \nodes \setminus \De[\graph]{n} \mid \Pa[\graph]{n} \quad \forall n \in \nodes \eq \label{eq:local_markov_property} \]
    Where $\indep$ denotes statistical independence. With this as motivation, we define a set of \term{causal parameters} for a causal structure $\graph$ to be a specification of a conditional distribution for every variable $n \in \nodes$ given its parents:
    \[ \bc{\prob[n \mid \Pa[\graph]{n}] \mid n \in \nodes} \eq \label{eq:causal_parameters} \]
    It is important to note that a set of causal parameters is not unique; the numerical nature of set of causal parameters is a \text{choice}. Altogether, a marginal model $\prob^{\mscenario}$ is \term{compatible} with a causal structure $\graph$ if there exists a choice of causal parameters such that each context $\prob[V] \in \prob^{\mscenario}$ can be \textit{recovered} from the following series of operations:
    \begin{enumerate}
        \item Obtain a joint probability distribution over \textit{all} variables of the causal structure using the chosen casual parameters:\footnote{Here $\prod$ denotes the product distribution $\prod_{n \in \nodes} \prob[n \mid \Pa[\graph]{n}] = \prob[n_1 \mid \Pa[\graph]{n_1}] \times \cdots \times \prob[n_k \mid \Pa[\graph]{n_k}]$.}
        \[ \prob[\nodes] = \prod_{n \in \nodes} \prob[n \mid \Pa[\graph]{n}] \eq \label{eq:local_markov_property_operation}\]
        This step ensures that the local Markov property (\cref{eq:causal_parameters}) holds for all variables $n \in \nodes$.
        \item Marginalize over the latent variables of $\graph$ to obtain a joint distribution over the observed variables $\nodes_O$:
        \[ \prob[\nodes_O] = \sum_{\nodes_L} \prob[\nodes] \eq \label{eq:observable_marginalization} \]
        \item Finally marginalize over the observed nodes not in $V$ to obtain $\prob[V]$:
        \[ \prob[V] = \sum_{\nodes_O \setminus V} \prob[\nodes_O] \eq \label{eq:context_marginalization} \]
    \end{enumerate}
    It is important to note that compatibility between $\prob^{\mscenario}$ and $\graph$ requires that it is possible to recover \textit{each} context $\prob[V] \in \prob^{\mscenario}$ from the \textit{same} set of causal parameters. If it is not possible for $\prob^{\mscenario}$ to be recovered for \textit{any} choice of causal parameters, then $\prob^{\mscenario}$ is \term{incompatible}.

    \subsection{Inequalities}
    As defined in \cref{sec:compatibility}, determining compatibility is equivalent to determining if a set of causal parameters exists with the above-mentioned properties. In practice however, checking causal parameters systemically is intractable, especially when the cardinality (number of outcomes) of the latent variables is unspecified. Instead, \term{causal compatibility inequalities}\footnote{We refer to these inequalities as ``causal compatibility inequalities'' instead of simply Bell inequalities for two reasons. First, ``Bell inequalities'' usually are associated specifically with the Bell scenario/Bell causal structure. Second, the inequalities derived in this work are fundamentally distinct from a typical Bell inequality in that these inequalities are \textit{polynomial} over $\prob^{\mscenario}$ instead of \textit{linear}.} are probabilistic inequalities defined over the marginal contexts $\prob[V] \in \prob^{\mscenario}$ within the marginal model. If a marginal model $\prob^{\mscenario}$ happens to \textit{violate} any causal compatibility inequality, then that marginal model is definitely \textit{incompatible}. In \cref{sec:deriving_inequalities}, a brief discussion of existing techniques for deriving these inequalities will take place.

@@

    Following the definition of causal compatibility from \cref{sec:causal_compatibility}, a distribution $\prob[\nodes_{O}] = \prob[ABC]$ is compatible with the triangle scenario if and only if there exists a choice of causal parameters $\bc{\prob[A|X,Y],\prob[B|Y,Z],\prob[C|Z,X],\prob[X],\prob[Y],\prob[Z]}$ such that the joint distribution $\prob[\nodes] = \prob[ABCXYZ]$ is a product over the causal parameters,
    \[ \prob[ABCXYZ] = \prob[A|X,Y]\prob[B|Y,Z]\prob[C|Z,X]\prob[X]\prob[Y]\prob[Z] \]
    And also, the given distribution $\prob[ABC]$ is a marginalization of $\prob[ABCXYZ]$ over $X, Y, Z$:
    \[ \prob[ABC] = \sum_{X,Y,Z}\prob[A|X,Y]\prob[B|Y,Z]\prob[C|Z,X]\prob[X]\prob[Y]\prob[Z] \eq \label{eq:triangle_compatibility} \]

    For certain causal structures, the conditional independence relations provided by $d$-separation relations are a sufficient characterization of compatibility. Recently, \citet{Henson_2014}, classified the triangle scenario as an \textit{interesting} causal structure in the sense that conditional independence relations \textit{do not} provide a sufficient characterization of non-classicality.\footnote{Note that the triangle scenario possesses \textit{zero} conditional independence relations over the observable nodes $A, B, C$.} Moreover, \citet{Inflation} concretely demonstrate that the space classical distributions on the triangle scenario is non-convex, unlike the Bell scenario. It can be argued that convexity of the Bell scenario distributions is partially responsible for the wealth of knowledge that is known about the Bell scenario. It is also reasonable to speculate that quantum non-classicality in the triangle can be translated directly into a computational advantage for certain computational circuits~\cite{Terhal_2002}. Characterizing quantum non-classicality for interesting causal structures remains an unsolved problem, and the triangle scenario is a small, non-trivial case study.

    Additionally, \citet{Fritz_2012} demonstrated that the triangle scenario is the \textit{smallest} correlation scenario\footnote{Correlation scenarios differ from the Bell scenario in that measure setting variables are non-existence in correlation scenarios.} in which their exists incompatible, quantum distributions by explicitly mentioning one which we have elected to call the Fritz distribution (see \cref{sec:fritz_distribution}). Fritz's proof of non-classicality does not involve compatibility inequalities, but instead mentions that inequalities would be helpful for characterizing non-classical in the triangle scenario.

>>>>

\PassOptionsToPackage{justification=justified}{caption}
% === Revtex Declaration ===
\documentclass[aps, 10pt, english, twoside, pra, nofootinbib, tightenlines, longbibliography]{revtex4-1}

@@

\begin{document}
    \title{Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Scenario}
    \author{Thomas C. Fraser}
    \email{tcfraser@tcfraser.com}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada \\ University of Waterloo, Waterloo, Ontario, Canada}
    % \author{Elie Wolfe}
    % \email{ewolfe@perimeter@institute.ca}
    % \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada}
    \date{\today}
    \begin{abstract}
        It has long been recognized that certain quantum correlations are often incompatible with particular assumption about classical causal structure. Such correlations are sometimes termed nonlocal, although it would be more accurate to say simply noncassical, as the presence of correlations certifies the genuinely nonclassical nature of a causal setup in a device-independent fashion.  Whenever a  quantum correlation violates a causal compatibility inequality (e.g. a Bell inequality) then the correlation is witnessed as nonclassical. Any constraint which is satisfied by all correlations which \emph{are} classically compatible with a given causal scenario defined a causal compatibility criterion. Such criteria were recently were recently derived for the triangle scenario [\href[pdfnewwindow]{http://arxiv.org/abs/1609.00672}{arXiv:1609.00672}] in the form of polynomial inequalities, begging the question: do any of those inequalities admit violation by quantum correlations? Numerical investigation suggests not, and we further conjecture that the set of correlations admitted by the classical triangle triangle scenario is equivalent to the set of correlations admitted by its quantum generalization whenever the three observable variables are binary. Our main contribution inn this work, however, is the derivation of new causal compatibility inequalities for the triangle scenario which \emph{do} admit quantum violation. We conclude by considering the possibility of quantum resources potentially qualitatively different from those known previously.
    \end{abstract}
    \maketitle
    \tableofcontents
    \setlength{\parskip}{1em}
    \clearpage

    \section{Introduction}
    \label{sec:introduction}
    In recent decades, the technical utility of quantum mechanics has become abundantly clear. In the realm of computation, quantum algorithms -- such as Shor's algorithm~\cite{Shor_1997} and numerous others~\cite{Jordan_2016} -- scale exponentially better than their classical counterparts. In the realm of secure communication, quantum protocols --- a popular example being Quantum Key Distribution~\cite{Bennett_2014} -- are able to provide privacy even against hypothetical adversaries with unlimited computational power, a desideratum which classical protocols are unable to fulfill. History, every opportunity of quantum phenomena which fails to be emulated by classical resources has led to a practical exploitation of the nonclassical feature in order to solve a some computational of communicational problem~\cite{Neilsen_Chaung_2011}. Motivated by past successes, a primal objective of modern quantum information theory is the discovery of new situations wherein quantum mechanics offers an advantage, and certification that the quantum advantage is genuine.

    From a foundational prospective, the most robust demonstrations of quantum phenemena with no classical emulation have involved Bell inequalities~\cite{Bell_1964}. Originally, Bell inequalities were derived as a way to show that no hidden variable theory could ever account for quantum mechanics; in this sense Bell inequalities are a response to the famous EPR paradox~\cite{EPR_Orig}. The enumeration of Bell inequalities has since become a widespread systematic way of demonstration the \textit{nonclassicality} of a given observation. More recently it has been appreciated that Bell inequalities can understood as consequences of \term{causal inference}~\cite{Wood_2012}. Causal inference is concerned with classifying observations into those which can and cannot be explained by a hypothesized causal structure. The abstract nature of causal inference is responsible for its presence in numerous scientific fields including machine learning and biology~\cite{Pearl_2009,Pearl_2009_tr}. Causal compatability inequalities, such as Bell of Instrumental inequalities, charechterize the space of observations that are compatible with a hypothesized causal structure, albeit the charechterization offered by practically deriveable inequalities often only an approximation. To derive the traditional Bell inequalities from causal inference one starts with a (classical) causal structure known as the Bell scenario, as depicted in \cref{fig:bell_scenario}. The fundamental Bell scenario involves noncommunicating parties making measurements on some hidden shared resource $\la$, where the measuremement outcomes ($A, B$) are presumed to be stochastic functions of the the local choices of measurement settings ($S_A, S_B$) and the shared resource $\la$. Quantum nonclassicality in the Bell scenario has been thoroughly studied since Bell's original work~\cite{Brunner_2013}. More complex scenarios, however, such as the correlation scenarios proposed by Fritz~\cite{Fritz_2012,Fritz_2014}, are much less understood. Here we investigate one particular causal scenario: the triangle scenario (\cref{fig:triangle_scenario}).

    The triangle scenario (\cref{fig:triangle_scenario}) is a causal structure composed of $3$ parties labeled $A, B, C$ arranged in a triangular configuration while pair-wise sharing hidden/latent variables $X, Y, Z$. It has been extensively studied previously (see~\cite[Fig. 1]{Steudel_2010},~\cite[Fig. 6]{Chaves_2014},~\cite[Fig. 8]{Branciard_2012},~\cite[Fig. 8, App. E]{Henson_2014},~\cite[Fig. 3]{Fritz_2012},~\cite[Fig. 1]{Inflation}, $\ldots$). An overview of some milestone results is provided in \cref{sec:triangle_scenario}. Identifying causal compatibility inequalities for this configuration has been seen as particularly challenging~\citet{Branciard_2012}. Further identifying causal compatibility inequalities of such high resolution that such inequalities can be violated by quantum-accessible distributions has remained out-of-reach for the triangle scenario.

    This work is the first, therefore, to find causal compatibility inequalities for the triangle scenario that are able to be violated by quantum-accessible distributions. This accomplishment was made possible through the combination of two previous developments. The first, an insight made by \citet{Fritz_2012} regarding the ability to re-interpret the Bell scenario as a portion of the triangle scenario. The second, a new framework for solving causal inference problems developed by Wolfe \emph{et al.}~\cite{Inflation} called \term{The Inflation Technique}. Ultimately, this work serves as a validation that the Inflation Technique is sensitive enough to offer new insights into quantum nonclassicality. Moreover, these new inequalities offer an avenue for recognizing previously unknown forms of nonclassicality. The authors attempts to find such novel resources were met with only partial success, suggesting clear refinements for future exploration.

    The manuscript is organized into the following sections: \Cref{sec:causal_compatibility} recalls important notions from causal inference theory and sets up the notation to be used. \Cref{sec:triangle_scenario} discusses the triangle scenario and provides an overview of existing research; identifying its stark differences from the Bell scenario and motivating why the triangle scenario was chosen as a case study. \Cref{sec:fritz_distribution} defines and discusses a singularly-quantum correlation first concieved of by \citet{Fritz_2012}, which we term the \term{Fritz distribution}. The Fritz distribution was proven to be non-classical without the use of inequalities --- this report fills in the desired missing inequalities. \Cref{sec:inflation_technique} briefly summaries the Inflation Technique in the specific context of this work. Although the summary presented here is designed to be self-standing, a much more pedagogical introduction is offered by the original work~\cite{Inflation}. \Cref{sec:deriving_inequalities} demonstrates how the inflation technique can be used to derive causal compatibility inequalities for the triangle scenario that are violated by quantum-accessible distributions. A sample of these inequalities are presented at the end of \cref{sec:deriving_inequalities} as well as in \cref{sec:web_inequality,sec:symmetic_web_inequality}. Finally, \cref{sec:violations_noise} introduces the method for parameterizing the space of quantum-accessible distributions which we used to numerically optimize against our derived inequalities in search of stronger types of non-classicality. Additionally, the Fritz distribution is subjected to noise in order to measure the robustness of the derived inequalities.

    \section{Causal Compatibility}
    \label{sec:causal_compatibility}
    The task of causal inference aims to determine the set of potentially observable probability distribution compatible with some hypothesis about causal relationships. whether or not a given family of observed probability distributions can be explained by a proposed causal mechanism~\cite{Pearl_2009}. If a set of observed distributions can be explained by the hypothesized causal mechanism, then the observations are said to be \textit{compatible} with the causal mechanism. In order to define compatibility rigorously, we first need to formally define what is meant by ``set of observations''.

    \subsection{Marginal Models}
    Given a set of observable variables $\jointvar = \bc{v_1, \ldots, v_k}$ where each $v \in \jointvar$ is a statistical random variable, one can conceivably make a measurement over of these variables and obtain a joint probability distribution $\prob[\jointvar] = \prob[v_1,\ldots,v_n]$. More generally however, and depending on the context, it is possible to make a measurement on a set of the system $V \subset \jointvar$ in order to obtain a marginal distribution $\prob[V]$. With complete generality however, one could in principle be limited so as to only be able to record observations on portions of the system $\mscenario = \bc{V_1, \ldots, V_m \mid \forall i : V_i \subseteq \jointvar}$ obtaining a collection of marginal probability distributions $\bc{\prob[V_1], \ldots, \prob[V_m] \mid \forall i: V_i \in \mscenario}$.

    The set $\mscenario = \bc{V \mid V \subseteq \jointvar}$ of subsets of $\jointvar$ is referred to as the \term{marginal scenario} and each element $V \in \mscenario$ is termed a \term{(marginal) context} of $\mscenario$. The complete set of marginal distributions is referred to as the \term{marginal model} and is denoted with an superscript $\prob^{\mscenario}$:
    \[ \prob^{\mscenario} = \bc{\prob[V] \mid V \in \mscenario} \]
    A marginal model acts as the most general description of a family of observations that can be made over $\jointvar$. Strictly speaking, as defined by~\cite{Fritz_2011}, a marginal scenario forms an \textit{abstract simplicial complex} where it is required that all subsets of contexts are also contexts.
    \[ \forall V \in \mscenario, V' \subset V : V' \in \mscenario \]
    Throughout this document, we exclusively consider (without loss of generality) the \textit{maximal} marginal scenario; restricting our focus to the largest contexts in the marginal scenario. Additionally, all marginal scenarios are taken to be \text{complete} in the sense that the marginal scenario covers the complete set of observable variables $\jointvar$.
    \[ \jointvar = \bigcup_{V \in \mscenario} V \]

    \subsection{Causal Structure}
    A hypothesis of causal mechanism is formally referred to as  a \term{causal structure}. Mathematically, a causal structure can be represented as a directed acyclic graph $\graph = \br{\nodes, \edges}$ of nodes $\nodes$ and edges $\edges$ as defined in \cref{sec:graph_theory}. The nodes of a causal structure are random variables and are classified into two disjoint classes: the \term{latent nodes} $\nodes_L$ representing the set of nodes that are unobservable and the \term{observed nodes} $\nodes_O$. The latent nodes of a causal structure represent completely hidden variables that are hidden by some fundamental process or cannot be measured due to other limitations. The observed nodes represent random variables that are measurable and in practice will be observed by the marginal model $\prob^\mscenario$. The edges $\edges$ of a causal structure represent casual influence between variables within the causal structure.

@@

    Now that the notions of a causal structure $\graph$ and a marginal model $\prob^{\mscenario}$ are defined, it is now possible to precisely define the notion of compatibility between the two. Intuitively, the absence of certain causal connections within $\graph$ should restrict and limit the set of possible observations that can be made on $\graph$; equivalently, the causal connections within $\graph$ should explain $\prob^{\mscenario}$. If each variable $n \in \nodes$ is caused by its parents $\Pa[\graph]{n}$, then $n$ should be independent of all other variables that are not its descendants. This is known as the \term{local Markov property}:
    \[ n \indep \nodes \setminus \De[\graph]{n} \mid \Pa[\graph]{n} \quad \forall n \in \nodes \eq \label{eq:local_markov_property} \]
    Where $\indep$ denotes statistical independence. With this as motivation, we define a set of \term{causal parameters} for a causal structure $\graph$ to be a specification of a conditional distribution for every variable $n \in \nodes$ given its parents:
    \[ \bc{\prob[n \mid \Pa[\graph]{n}] \mid n \in \nodes} \eq \label{eq:causal_parameters} \]
    It is important to note that a set of causal parameters is not unique; the numerical nature of set of causal parameters is a \text{choice}. Altogether, a marginal model $\prob^{\mscenario}$ is \term{compatible} with a causal structure $\graph$ if there exists a choice of causal parameters such that each context $\prob[V] \in \prob^{\mscenario}$ can be \textit{recovered} from the following series of operations:
    \begin{enumerate}
        \item Obtain a joint probability distribution over \textit{all} variables of the causal structure using the chosen casual parameters:\footnote{Here $\prod$ denotes the product distribution $\prod_{n \in \nodes} \prob[n \mid \Pa[\graph]{n}] = \prob[n_1 \mid \Pa[\graph]{n_1}] \times \cdots \times \prob[n_k \mid \Pa[\graph]{n_k}]$.}
        \[ \prob[\nodes] = \prod_{n \in \nodes} \prob[n \mid \Pa[\graph]{n}] \eq \label{eq:local_markov_property_operation}\]
        This step ensures that the local Markov property (\cref{eq:causal_parameters}) holds for all variables $n \in \nodes$.
        \item Marginalize over the latent variables of $\graph$ to obtain a joint distribution over the observed variables $\nodes_O$:
        \[ \prob[\nodes_O] = \sum_{\nodes_L} \prob[\nodes] \eq \label{eq:observable_marginalization} \]
        \item Finally marginalize over the observed nodes not in $V$ to obtain $\prob[V]$:
        \[ \prob[V] = \sum_{\nodes_O \setminus V} \prob[\nodes_O] \eq \label{eq:context_marginalization} \]
    \end{enumerate}
    It is important to note that compatibility between $\prob^{\mscenario}$ and $\graph$ requires that it is possible to recover \textit{each} context $\prob[V] \in \prob^{\mscenario}$ from the \textit{same} set of causal parameters. If it is not possible for $\prob^{\mscenario}$ to be recovered for \textit{any} choice of causal parameters, then $\prob^{\mscenario}$ is \term{incompatible}.

    \subsection{Inequalities}
    As defined in \cref{sec:compatibility}, determining compatibility is equivalent to determining if a set of causal parameters exists with the above-mentioned properties. In practice however, checking causal parameters systemically is intractable, especially when the cardinality (number of outcomes) of the latent variables is unspecified. Instead, \term{causal compatibility inequalities}\footnote{We refer to these inequalities as ``causal compatibility inequalities'' instead of simply Bell inequalities for two reasons. Firstly, ``Bell inequalities'' usually are associated specifically with the Bell scenario/Bell causal structure. Secondly, the inequalities derived in this work are fundamentally distinct from a typical Bell inequality in that these inequalities are \textit{polynomial} over $\prob^{\mscenario}$ instead of \textit{linear}.} are probabilistic inequalities defined over the marginal contexts $\prob[V] \in \prob^{\mscenario}$ within the marginal model. If a marginal model $\prob^{\mscenario}$ happens to \textit{violate} any causal compatibility inequality, then that marginal model is definitely \textit{incompatible}. In \cref{sec:deriving_inequalities}, a brief discussion of existing techniques for deriving these inequalities will take place.

@@

    Following the definition of causal compatibility from \cref{sec:causal_compatibility}, a distribution $\prob[\nodes_{O}] = \prob[ABC]$ is compatible with the triangle scenario if and only if there exists a choice of causal parameters $\bc{\prob[A|X,Y],\prob[B|Y,Z],\prob[C|Z,X],\prob[X],\prob[Y],\prob[Z]}$ such that 
    %the joint distribution $\prob[\nodes] = \prob[ABCXYZ]$ is a product over the causal parameters,
    %\[ \prob[ABCXYZ] = \prob[A|X,Y]\prob[B|Y,Z]\prob[C|Z,X]\prob[X]\prob[Y]\prob[Z] \]
    %And also, the given distribution 
    $\prob[ABC]$ is a marginalization of $\prob[ABCXYZ]$ over $X, Y, Z$:
    \[ \prob[ABC] = \sum_{X,Y,Z}\prob[A|X,Y]\prob[B|Y,Z]\prob[C|Z,X]\prob[X]\prob[Y]\prob[Z] \eq \label{eq:triangle_compatibility} \]

    For certain causal structures, the conditional independence relations provided by $d$-separation relations are a sufficient characterization of compatibility. Recently, \citet{Henson_2014}, classified the triangle scenario as an \textit{interesting} causal structure in the sense that conditional independence relations \textit{do not} provide a sufficient characterization of non-classicality.\footnote{Note that the triangle scenario possesses \textit{zero} conditional independence relations over the observable nodes $A, B, C$.} Moreover, \citet{Inflation} concretely demonstrate that the space of classical distributions on the triangle scenario is non-convex, unlike the Bell scenario. It can be argued that convexity of the Bell scenario distributions is partially responsible for the wealth of knowledge that is known about the Bell scenario. It is also reasonable to speculate that quantum non-classicality in the triangle can be translated directly into a computational advantage for certain computational circuits~\cite{Terhal_2002}. Characterizing quantum non-classicality for interesting causal structures remains an unsolved problem, and the triangle scenario is a small, non-trivial case study.

    Additionally, \citet{Fritz_2012} demonstrated that the triangle scenario is the \textit{smallest} correlation scenario\footnote{Correlation scenarios differ from the Bell scenario in that measure setting variables are non-existent in correlation scenarios.} in which their exists incompatible, quantum distributions by explicitly mentioning one which we have elected to call the Fritz distribution (see \cref{sec:fritz_distribution}). Fritz's proof of non-classicality does not involve compatibility inequalities, but instead mentions that inequalities would be helpful for characterizing non-classical in the triangle scenario.
     \comment[E]{Revisions processed up to here thusfar.}
